\normalfont{A bird's-eye view on the habitability of exoplanets via statistical learning techniques}

### 2020-10-01 Manuscript/Manuscript.tex
PUB

The first approach that was considered is the classification decision tree: as we can see from Fig \ref{DecisionTree} the main discriminant is the stellar luminosity: this feature discriminate,as one expects, the largest part of the habitable planet from the non-habitable. The nodes $4$, $9$ and $19$ discriminates deeply the habitability on the basis of stellar temperature planet temperature and its main distance from the host star. Note that the mean distance (P\_D) has a less importance: this can be ascribed to the fact that the habitability was extended beyond the "earth zone" by considering habitable planets that, with a particular atmosphere as shown in fig \ref{Planets_habitability_Seager}, have the correct surface temperature for liquid water. 
\footnote{In principle one may wonder if a better approach should considered only the planets like Earth habitable or perhaps to distinguish the two cases in order to get the distance effect: the issue here is that these exoplanets are very few, thus to preserve the significance of the habitable planets, if such approach was undertaken, a further reduction of the dataset had to be considered. Since this choice strongly limits the quality of forecasts, such approach was not considered.}. The performance of this algorithm on this dataset, evaluated with a 10-fold cross-validation, is reported in Fig \ref{Decision_tree_performances} and in the table \ref{DecisionTree_perf_tab}. Next the random forest approach was considered.First consider the plot in Fig. \ref{Random_forest_trees}: at first when the number of trees is moved is increased the error for

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{Pic/DecisionTree.pdf}
\caption{The classification of exoplanets habitability obtained by the decision tree algorithm (1 stands for habitable while 0 not habitable). The highest node $1$ involves the Stellar Luminosity (S\_L), then follows the nodes $2$ and $4$ that involve the stellar temperature (S\_T), and finally the $9$ and $19$ that involve the planet temperature equilibrium (P\_T\_E) and the mean planet distance (P\_D). The units of measure are reported in the table \ref{features_table}. The highest number in each boxc refers to the output of the classifier give the condition reported on his bottom, the two number in the middle the percentages of the splitting and the last number the overall percentage of the planes classified by the node. Note that, since only boxes for which more than 5 items were considered, the overall sum is lower than $100 \%$}
\label{DecisionTree}
\end{center}
\end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{Pic/Random_forest_trees.pdf}
\caption{}
\label{Random_forest_trees}
\end{center}
\end{figure}

>>>>

The first approach that was considered is the classification decision tree: as we can see from Fig \ref{DecisionTree} the main discriminant is the stellar luminosity: this feature discriminate,as one expects, the largest part of the habitable planet from the non-habitable. The nodes $4$, $9$ and $19$ discriminates deeply the habitability on the basis of stellar temperature planet temperature and its main distance from the host star. Note that the mean distance (P\_D) has a less importance: this can be ascribed to the fact that the habitability was extended beyond the "earth zone" by considering habitable planets that, with a particular atmosphere as shown in fig \ref{Planets_habitability_Seager}, have the correct surface temperature for liquid water. 
\footnote{In principle one may wonder if a better approach should considered only the planets like Earth habitable or perhaps to distinguish the two cases in order to get the distance effect: the issue here is that these exoplanets are very few, thus to preserve the significance of the habitable planets, if such approach was undertaken, a further reduction of the dataset had to be considered. Since this choice strongly limits the quality of forecasts, such approach was not considered.}. The performance of this algorithm on this dataset, evaluated with a 10-fold cross-validation, is reported in Fig \ref{Decision_tree_performances} and in the table \ref{DecisionTree_perf_tab}. Next the random forest approach was considered. As shown in Fig. \ref{Random_forest_trees} the main parameters for this parameters were previously checked. The results of the random forest algorithm are reported in Fig

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{Pic/cp_vs_ROC.pdf}
\caption{Perfomance of the decision tree as the cp parameter grows evaluated with the ROC metric. }
\label{cp_vs_ROC}
\end{center}
\end{figure}

\begin{figure}[h]
 \centering
\begin{center}
\subfloat[]{\includegraphics[width=0.8\textwidth]{Pic/DecisionTree.pdf}}\\
\subfloat[]{\includegraphics[width=0.8\textwidth]{Pic/Var_imp_dec_tree.pdf}}
\caption{The classification of exoplanets habitability obtained by the decision tree algorithm (1 stands for habitable while 0 not habitable). The units of measure are reported in the table \ref{features_table}. In the  upper panel (a) the decision tree representation is plotted: the highest number in each box refers to the output of the classifier give the condition reported on his bottom, the two number in the middle the percentages of the splitting and the last number the overall percentage of the planes classified by the node. Note that, since only boxes for which more than 5 items were considered, the overall sum is lower than $100 \%$. In the lower panel (b) is reported the overall importance of each variable.}
\label{DecisionTree}
\end{center}
\end{figure}

@@

\begin{figure}[h]
  \centering
\subfloat[]{\includegraphics[width=0.5\textwidth]{Pic/Random_forest_trees.pdf}}
\subfloat[]{\includegraphics[width=0.5\textwidth]{Pic/mtry_Random_forest.pdf}}
\caption{The performances of the random forest algorithm: as the ensemble grew (panel a) accompanied by the optimization of the number of variables available (mtry). For the calculations presented in this work an ensemble of 500 trees (the default one) was considered while the mytry parameter was set equal to 3.}
\label{Random_forest_trees}
